---
title: "Prepare aquaculture nodes"
author: "Simone Stevenson"
date: "17/05/2022"
output: html_document
objective: To prepare the nodes of New Zealand's aquaculture traffic network
---

```{r clear space,include = FALSE}

rm(list = ls())

```


```{r setup, include=FALSE}

# https://bookdown.org/yihui/rmarkdown/r-code.html

knitr::opts_chunk$set(echo = FALSE, message = FALSE,
  warnings = FALSE, error = FALSE)

```

```{r project info}

# Projection info

NumericEPSG <- 3994

CentreCoords <- c(-42.22, 173.40)

# These will be used to label your output files, update at beginning of session

pathway <- "aquaculture"

Component <- "Nodes"

#If you need to add an additional descriptor to your file (if not just make NA)
Label <- "OnWaterFarms"

# IMPORTANT - IF YOU ARE MAKING A NEW VERSION UPDATE THIS NUMBER OR YOU WILL OVERWRITE PREV VERSIONS
Version <- "0"

Author <- "SS"

CleanFileName <- paste0("Clean_", pathway, Component, "_", Author, "_", Version)

CleanFileName

WIPFileName <- paste0("WIP_", pathway, Component, "_", Author, "_", Version)

WIPFileName

```
Directory path to git repo

Simone's deakin laptop

cd "C:/Users/ssimone/OneDrive - Deakin University/MSEAC/Code/git_code/RA1.4_Aquaculture"

remote repo https://github.com/Simonestevo/RA1.4_Aquaculture

```{r libraries}
# Install packages as needed (will install any packages you don't already have)
if (!require(tidyverse)) install.packages('tidyverse')
if (!require(tidylog)) install.packages('tidylog')
if (!require(sf)) install.packages('sf')
if (!require(leaflet)) install.packages('leaflet')
if (!require(scales)) install.packages('scales')

## Data wrangling
library(tidyverse)
library(tidylog)
library(sf)

## Analysis


## Viz
library(leaflet)
library(scales)

```


```{r ggplot theme}

plot_theme <- theme(panel.background = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              legend.position = "bottom",
              legend.key.size = unit(0.25, 'cm'),
              plot.title = element_text(size = 10, face = "bold"),
              legend.title=element_text(size=6), 
              legend.text=element_text(size=5),
              axis.line = element_line(colour = "black"),
              axis.title.x = element_text(size=8),
              axis.title.y = element_text(size=8),
              axis.text.x = element_text(size=8),
              axis.text.y = element_text(size=8),
              axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())

```

```{r colour palettes, eval=FALSE}

show_col(viridis_pal()(20))
show_col(viridis_pal(option = "plasma")(20))
show_col(viridis_pal(option = "magma")(20))
show_col(viridis_pal(option = "inferno")(20))
show_col(viridis_pal(option = "mako")(20))
show_col(viridis_pal(option = "cividis")(20))
show_col(viridis_pal(option = "rocket")(20))
show_col(viridis_pal(option = "turbo")(20))

```

```{r set up directories}

ProjectDirectory <- 'S:/RDS28221-BiosecurityToolbox/projectData'

# Get the date

today <- Sys.Date()

# Create a directory to store raw inputs as received

inputs <- file.path(ProjectDirectory, "1_rawInputs")

if (!"1_rawInputs" %in% basename(list.dirs(ProjectDirectory))) { # Check folder has not already been created
  
  dir.create(inputs)
  
  print("creating directory to store input files")
  
} else {
  
  print("found existing input directory")
  
}

# Create a directory to store current work in progress and files created in R

wipR <- file.path(ProjectDirectory, "2_wipR")

if (!"2_wipR" %in% basename(list.dirs(ProjectDirectory))) { # Check folder has not already been created
  
  dir.create(wipR)
  
  print("creating directory to store R wip files")
  
} else {
  
  print("found existing wip directory")
  
}

# Create a directory to store current work in progress and files created in GIS
# software then used by this code

wipGIS <- file.path(ProjectDirectory, "3_wipGIS")

if (!"3_wipGIS" %in% basename(list.dirs(ProjectDirectory))) { # Check folder has not already been created
  
  dir.create(wipGIS)
  
  print("creating directory to store GIS wip files")
  
} else {
  
  print("found existing GIS wip directory")
  
}

# Create a directory to store all finalised outputs

cleanOutputs <- file.path(ProjectDirectory, "4_cleanOutputs")

if (!"4_cleanOutputs" %in% basename(list.dirs(ProjectDirectory))) { # Check folder has not already been created
  
  dir.create(cleanOutputs)
  
  print("creating directory to store outputs")
  
} else {
  
  print("found existing outputs directory")
  
}

```


```{r load data, include = FALSE}
list.files(file.path(inputs, pathway))

FarmsShp <- st_read(file.path(inputs, pathway, "MPICurrentMarineFarms",
                              "MPI_CurrentMarineFarms_11Dec2020.shp"))

list.files(file.path(inputs, "geographic"), recursive = TRUE)

RegionsShp <- st_read(file.path(inputs, "geographic", "statsnzregional-council-2020-generalised-SHP",
                              "regional-council-2020-generalised.shp"))
```

```{r look at data}

# Check the projection
st_crs(FarmsShp)

# Check attributes and for NA vals
summary(FarmsShp)

# Remove the Z dimension (depth??)

FarmsShp <- st_zm(FarmsShp)

# Project into WGS84 for the leaflet map
FarmsShpWGS84 <- st_transform(FarmsShp, 4326)
RegionsShpWGS84 <- st_transform(RegionsShp, 4326) %>% 
                   mutate(col = "green")

RegionPal <- colorFactor("viridis", domain = as.factor(RegionsShpWGS84$REGC2020_1))

# Check it looks right

st_crs(FarmsShpWGS84)

# Make a quick plot to look at the data
FarmsShpWGS84 %>% 
  leaflet() %>% 
  addTiles(group = "Basemap") %>% 
  addPolygons(data = RegionsShpWGS84, weight=2,col = 'black',
              fillColor = ~RegionPal(REGC2020_1),
              highlightOptions = highlightOptions(color='white',weight=1,
                                                  bringToFront = TRUE)) %>% 
    addPolygons(group = "Client_nam") %>% 
  setView(173.40, -42.22, zoom = 5)

```

```{r get centroids}

FarmsShp <- FarmsShp %>% 
            mutate(Centroids = st_centroid(geometry)) 

# Reproject the centroids so we can look on leaflet map
CentroidsWGS84 <- st_transform(FarmsShp$Centroids, 4326)

# Make a quick plot to look at the data
FarmsShpWGS84 %>% 
  leaflet() %>% 
  addTiles(group = "Basemap") %>% 
  addPolygons(group = "Client_nam") %>% 
  addCircles(data = CentroidsWGS84, color = "red") %>% 
  setView(173.40, -42.22, zoom = 5)

```

Make a shp file that matches the other nodes, with the following attributes:

FID = ID number
Shape = Point
Region = Region name (string)
src = Source of the data (string)
Notes = can leave empty
Name = Farm name (string)
Type = Aquaculture (string)
CLASS = Farm type?
Lat = geom
Lon = geom
Species = species or taxa name (string)

```{r wrangle data to match other nodes}

FarmsPoints <- FarmsShp %>% 
               dplyr::select(Client_nam, Species_Gr, Centroids) %>% 
               st_drop_geometry() %>% 
               mutate(SRC = "MPI",
                      Notes = NA,
                      Type = "Aquaculture") %>% 
               rename(geometry = Centroids,
                      Name = Client_nam,
                      Class = Species_Gr) 
               
head(FarmsPoints)
FarmsPoints <- FarmsPoints %>% 
               st_as_sf(crs = st_crs(FarmsShp)) %>% 
              mutate(Lon = sf::st_coordinates(.)[,1],
                Lat = sf::st_coordinates(.)[,2])

RegionsShp <- st_transform(RegionsShp, crs = st_crs(FarmsPoints)) 

RegionsShp <- RegionsShp %>% 
              dplyr::select(REGC2020_1, geometry) %>% 
              rename(Region = REGC2020_1)

# Add in regions

FarmsPoints <- FarmsPoints %>% 
               st_join(RegionsShp, join = st_intersects) %>% 
               dplyr::select(Region, SRC, Notes, Name, Type, Class, Lat, Lon,
                             geometry)

head(FarmsPoints)

class(FarmsPoints)

plot(st_geometry(FarmsPoints))

```

```{r save wip outputs}

# We are still missing the on shore nodes which need to be added to the final clean
# dataset, but the farm only data can be saved in the wip outputs if needed

# st_write(FarmsPoints, file.path(wipR, paste0(WIPFileName, ".shp")))

```

```{r save clean outputs}

# When final outputs are ready save them in the clean outputs folder

# st_write(FarmsPoints, file.path(cleanOutputs, paste0(WIPFileName, ".shp")))

```

